---
title: "African Institute for Mathematical Sciences, Rwanda"
subtitle: "Bash/R Programming Tutorials"
output: pdf_document
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Requirements

-   Group work to solve the following questions

-   Each group will present

## Questions

a.  Import the `data.xlsx` data in R. Install the `openxlsx` package if you do not have it installed.

```{r warning=FALSE, message=FALSE}
library(openxlsx) # use function read.xlsx
# tuto_data <- complete the code
tuto_data <- read.xlsx("./data.xlsx")
tuto_data
```

b.  Calculate the percentage of missing values in each column

```{r}
is.na(tuto_data)
summary(tuto_data)
37/length(tuto_data$X1)*100
25/length(tuto_data$X2)*100
18/length(tuto_data$X3)*100
45/length(tuto_data$X4)*100
```

c.  How many observations will we get if we were to remove each row containing missing values?

```{r}
boxplot(tuto_data$X1)
boxplot(tuto_data$X2)
boxplot(tuto_data$X3)
boxplot(tuto_data$X4)


```

d.  Replace NAs in each column

    -   by its average if the column is numeric
    -   by its mode if the column is categorical

e.  Draw the box plots of numeric column

```{r}

```

f.  Make a scatter plot of `X3` and `X4` colored by `Class`. What do you observe?

```{r}
plot(X3~X4, data = tuto_data)
```

g.  What is the percentage of each class in the NA-free (imputed) data

```{r}

```

h.  What is the average of `X2` in each `Class`?
```{r}
mean(tuto_data$X2, na.rm = TRUE)
```

